---
title: "Immune Infiltration"
output:
  html_notebook: default
---

# Purpose

To see how many enzymes making use of metabolites in the study are targetable by drugs: 

```{r knitrSetup, include=FALSE}
library(knitr)
opts_chunk$set(out.extra='style="display:block; margin: auto"', fig.align="center", tidy=FALSE)
```

# Setup

```{r loadPackages}
library(xlsx)
```

# Read input files

```{r}
rnaSeqImmuneDecov <- read.xlsx2("../data/Metabolomics_ssGSEA.xlsx", "RNA-Seq Immune Deconvolution", startRow=2, stringsAsFactors=FALSE)
tmpNormData <- read.xlsx2("../data/Metabolomics_ssGSEA.xlsx", "ResuspensionNormData", header=FALSE, stringsAsFactors=FALSE)
```

# Extract identifier information from norm data
```{r}
tmp <- tmpNormData[1:12, 13:61]
samples <- tmp[3, 2:ncol(tmp)]
```

# Extract data and replace the column appropriate headers with the identifiers 
```{r}
tmp <- tmpNormData[14:nrow(tmpNormData), ]
colnames(tmp) <- tmpNormData[13,]
colnames(tmp)[14:ncol(tmp)] <- samples 
normData <- tmp 
```

# Compare identifiers in the two worksheets 
```{r}
idx <- which(samples %in% as.character(rnaSeqImmuneDecov$Sample))
selectedSamples <- as.character(samples)
selectedSamples <- selectedSamples[idx]

dropSamples <- c("JHNK474", "JHNK522") 
selectedSamples <- selectedSamples[!(selectedSamples %in% dropSamples)]
```

# Provide rownames for subsetting
```{r}
rownames(rnaSeqImmuneDecov) <- rnaSeqImmuneDecov$Sample
```

# Calculate correlations for each metabolite in the norm sheet to the scores in the deconv sheet
```{r, warning=FALSE}
rows <- 1:nrow(normData)
#cols <- 3:ncol(rnaSeqImmuneDecov)
cols <- 3:32

cors <- NULL
pvals <- NULL 

pb <- txtProgressBar(min=1, max=max(rows), style=3)

for(i in rows) {
  #setTxtProgressBar(pb, i)
  
  for(j in cols) {

    #i <- 1
    #j <- 3
    x <- normData[i, selectedSamples]
    y <- rnaSeqImmuneDecov[selectedSamples, j]
    names(y) <- selectedSamples
    
    x <- as.numeric(x)
    y <- as.numeric(y)
    
    t1 <- cor.test(x, y, method="spearman")
    cors <- c(cors, t1$estimate)
    pvals <- c(pvals, t1$p.value)
  }
}
```

# Adjust p-values 
```{r}
pvalsAdj <- p.adjust(pvals, method="fdr")

summary(pvals)
summary(pvalsAdj)
```

# Make matrix, ensure values from vector become rows 
```{r}
corsMat <- matrix(cors, ncol=length(cols), byrow=TRUE)
pvalsAdjMat <- matrix(pvalsAdj, ncol=length(cols), byrow=TRUE)
pvalsMat <- matrix(pvals, ncol=length(cols), byrow=TRUE)

colnames(corsMat) <- colnames(rnaSeqImmuneDecov)[cols]
colnames(pvalsAdjMat) <- colnames(rnaSeqImmuneDecov)[cols]
colnames(pvalsMat) <- colnames(rnaSeqImmuneDecov)[cols]

rownames(corsMat) <- normData$BIOCHEMICAL
rownames(pvalsAdjMat) <- normData$BIOCHEMICAL
rownames(pvalsMat) <- normData$BIOCHEMICAL
```

# Extract out significant values and save to files
```{r}
pvalThreshold <- 0.05

# ALL 
metabIdx <- NULL 

for(i in 1:nrow(corsMat)) {
  tmp <- which(as.numeric(pvalsMat[i,]) < pvalThreshold)
  if(length(tmp) > 0) {
    metabIdx <- c(metabIdx, i)
  }
}

## Rows meeting significance threshold
length(metabIdx)

write.csv(corsMat[metabIdx, ], "../results/immuneInfiltrationAllColsCors.csv")
write.csv(pvalsAdjMat[metabIdx, ], "../results/immuneInfiltrationAllColsPValsAdj.csv")
write.csv(pvalsMat[metabIdx, ], "../results/immuneInfiltrationAllColsPVals.csv")


# Score Columns 
metabIdx <- NULL
scoreCols <- 1:3

for(i in 1:nrow(corsMat)) {
  tmp <- which(as.numeric(pvalsAdjMat[i,scoreCols]) < pvalThreshold)
  if(length(tmp) > 0) {
    metabIdx <- c(metabIdx, i)
  }
}

## Rows meeting significance threshold
length(metabIdx)

# tx3 <- corsMat[metabIdx, scoreCols]
# write.csv(tx3, "../results/3Cols.csv")
```

# Check "StromalScore" (Column 3) for "ornithine"
```{r}
idx <- which( normData$BIOCHEMICAL == "ornithine")
x <- normData[idx, selectedSamples]
y <- rnaSeqImmuneDecov[selectedSamples, 3]

names(y) <- selectedSamples

x <- as.numeric(x)
y <- as.numeric(y)

plot(x, y)
```

# Heatmap 
```{r}
# nba <- read.csv("http://datasets.flowingdata.com/ppg2008.csv")
# nba$Name <- with(nba, reorder(Name, PTS))
# 
# nba.m <- melt(nba)
# nba.m <- ddply(nba.m, .(variable), transform, rescale = scale(value))
# 
# base_size <- 9
# 
# p <- ggplot(nba.m, aes(variable, Name)) + 
#   geom_tile(aes(fill = rescale), colour = "white") + 
#   scale_fill_gradient(low = "white", high = "steelblue")
# 
# p <- p + theme_grey(base_size = base_size) + 
#   labs(x = "", y = "") + 
#   scale_x_discrete(expand = c(0, 0)) +
#   scale_y_discrete(expand = c(0, 0)) + 
#   theme(legend.position = "none", axis.ticks = element_blank()) + 
#   ggtitle("Title")
# 
# p 
```