---
title: "Pharmacology Coverage Analysis"
output:
  html_notebook: default
---

## Purpose

To see how many enzymes making use of metabolites in the study are targetable by drugs: 

```{r knitrSetup, include=FALSE}
library(knitr)
opts_chunk$set(out.extra='style="display:block; margin: auto"', fig.align="center", tidy=FALSE)
```

# Setup

```{r loadPackages}
library(paxtoolsr)
library(data.table)

verbose <- TRUE
```

## Load Data

```{r loadData}
keggSif <- downloadPc2("PathwayCommons.8.kegg.EXTENDED_BINARY_SIF.hgnc.txt.gz", version = "8")
keggNodes <- keggSif$nodes

dbSif <- downloadPc2("PathwayCommons.8.drugbank.EXTENDED_BINARY_SIF.hgnc.txt.gz", version = "8")
dbNodes <- dbSif$nodes
```

## Load Local Data

```{r loadLocalData}
# KEGG IDs for pathways
load(file.path("..", "data", "KEGGpathays.RData"))

# KEGG IDs found in dataset
keggIds <- read.table(file.path("..", "import", "tempdir", "KEGG_dictionary.csv"), sep=",", header=TRUE, stringsAsFactors = FALSE)

chebiIds <- read.table(file.path("..", "import", "tempdir", "ChEBI_dictionary.csv"), sep=",", header=TRUE, stringsAsFactors = FALSE)

idx <- which(chebiIds$X %in% keggNodes$PARTICIPANT)
length(idx)
```

# Pharmacology
## Get genes in KEGG
```{r}
tmp <- chebiIds$X[idx]
intCats <- getSifInteractionCategories()

# Selected PC interactions
selIntCats <- c("consumption-controlled-by", "controls-production-of", "controls-transport-of-chemical")
#selIntCats <- c("consumption-controlled-by")

#keggSmMolSif <- filterSif(keggSif$edges, ids=tmp, interactionTypes=intCats$BetweenProteinSmallMolecule)
keggSmMolSif <- filterSif(keggSif$edges, ids=tmp, interactionTypes=selIntCats)
```

## Get interactions in KEGG that involve genes present in both KEGG and DrugBank
```{r}
# Get gene set for KEGG interactions between proteins and small molecules (as opposed to between two proteins)
x1 <- convertSifToGmt(keggSmMolSif, "db", returnSmallMolecules=FALSE)

# Get the interactions from DrugBank based on genes from KEGG 
tmpNodes <- dbNodes$PARTICIPANT
x2 <- tmpNodes[tmpNodes %in% x1$db]
x3 <- sort(unique(x2))
#x4 <- filterSif(dbSif, ids=x3, idsBothParticipants = FALSE)

# Get consumption-controlled-by interactions in KEGG that involve genes present in both KEGG and DrugBank
# NOTE: consumption-controlled-by interactions are those that are like
keggDbSif <- filterSif(keggSmMolSif, ids=x3, interactionTypes = selIntCats, idsBothParticipants = FALSE)
```

## Load Data on the Groups and Categories of DrugBank Data
```{r}
# Uses DrugBank: 5.0.3
# Find SIF CHEBI IDs that are in the DrugBank categories table
drugbankGroupsCategories <- readRDS(file.path("..", "data", "pharmacology_drugbank", "drugbankGroupsCategories.rds"))
#dbIdx <- which(keggDbSif$PARTICIPANT_A %in% drugbankGroupsCategories$PARTICIPANT)
#dbOut <- drugbankGroupsCategories[dbIdx,]
```

## Merge KEGG and DrugBank Datasets and Filter Out Particular Entries
```{r}
# (10) KEGG_CHEBI, KEGG_INTERACTION_TYPE, GENE, DRUGBANK_INTERACTION_TYPE, DRUGBANK_CHEBI, KEGG_NAME, DRUGBANK_NAME, DRUGBANK_CATEGORY, DRUGBANK_ID, KEGG_PATHWAY_NAMES
# (8) KEGG_CHEBI, GENE, DRUGBANK_CHEBI, KEGG_NAME, DRUGBANK_NAME, DRUGBANK_CATEGORY, DRUGBANK_ID, KEGG_PATHWAY_NAMES

nrow(keggSmMolSif)

k1 <- keggSmMolSif
drops <- c("INTERACTION_PUBMED_ID", "INTERACTION_DATA_SOURCE", "INTERACTION_TYPE", "MEDIATOR_IDS", "class")
setDF(k1)
k1 <- k1[, !(colnames(k1) %in% drops)]

b1 <- merge(k1, dbSif$edges, by.x="PARTICIPANT_B", by.y="PARTICIPANT_B", all.x=TRUE)
drops <- c("INTERACTION_PUBMED_ID", "INTERACTION_DATA_SOURCE", "INTERACTION_TYPE", "MEDIATOR_IDS", "PATHWAY_NAMES.y", "class")
b1 <- b1[, !(colnames(b1) %in% drops)]
colnames(b1) <- c("GENE", "KEGG_CHEBI", "KEGG_PATHWAY_NAMES", "DRUGBANK_CHEBI")
nrow(b1)

c1 <- merge(b1, drugbankGroupsCategories, by.x="DRUGBANK_CHEBI", by.y="PARTICIPANT", all.x=TRUE)
c1$DRUGBANK_NAME <- c1$PARTICIPANT_NAME
drops <- c("PARTICIPANT_NAME")
c1 <- c1[, !(colnames(c1) %in% drops)]
nrow(c1)

d1 <- merge(c1, keggNodes, by.x="KEGG_CHEBI", by.y="PARTICIPANT", all.x=TRUE)
drops <- c("PARTICIPANT_TYPE", "UNIFICATION_XREF", "RELATIONSHIP_XREF")
d1 <- d1[, !(colnames(d1) %in% drops)]

d1$KEGG_NAME <- d1$PARTICIPANT_NAME
drops <- c("PARTICIPANT_NAME")
d1 <- d1[, !(colnames(d1) %in% drops)]
nrow(d1)

cols <- c("KEGG_CHEBI", "GENE", "DRUGBANK_CHEBI", "KEGG_NAME", "DRUGBANK_NAME", "DRUGBANK_CATEGORY", "DRUGBANK_GROUPS", "DRUGBANK_ID", "KEGG_PATHWAY_NAMES")
e1 <- d1[, cols]
  
# Exclude certain drug types of less interest
excludeGroups <- c("experimental", "illicit", "nutraceutical", "withdrawn")
f1 <- lapply(e1$DRUGBANK_GROUPS, function(x) {
  a <- all(!(x %in% excludeGroups))
  b <- length(x) > 0
  c <- !(is.na(x))
  r <- all(c(a, b, c))
})
f2 <- unlist(f1)
f4 <- e1[f2,]
nrow(f4)

# Include only certain drug categories of interest
j1 <- lapply(f4$DRUGBANK_CATEGORY, function(x) {
  a <- any(grepl("enzyme inhibit", x, ignore.case=TRUE))
  b <- any(grepl("antimetabolites", x, ignore.case=TRUE))
  r <- any(c(a, b))
})
j2 <- unlist(j1)
j4 <- f4[j2,]
nrow(j4)

# Exclude interactions that focus on drug metabolism because these do inhibit enzyme function
m1 <- lapply(j4$KEGG_PATHWAY_NAMES, function(x) {
  a <- any(!grepl("drug metabolism", x, ignore.case=TRUE))
  r <- any(c(a))
})
m2 <- unlist(m1)
m4 <- j4[m2,]
nrow(m4)

# Remove the generic "Metabolic pathways"
g2 <- lapply(m4$KEGG_PATHWAY_NAMES, function(x) {
  x[!(x %in% "Metabolic pathways")]  
})
m4$KEGG_PATHWAY_NAMES <- g2

f6 <- convertDataFrameListsToVectors(m4)
write.table(f6, file="kegg_drugbank_results.txt", sep="\t", quote=FALSE, row.names=FALSE)
```

## Get Pubmed References for the Remaining Interactions
```{r}
library(RISmed)
library(simpleRCache)

setCacheRootPath("cache")
EUtilsSummaryCached <- addMemoization(EUtilsSummary)
EUtilsGetCached <- addMemoization(EUtilsGet)

pmidCnts <- NULL 
pmids <- list() 

pb <- txtProgressBar(min=1, max=nrow(m4), style=3)

for(i in 1:nrow(m4)) {
  #i <- 1
  query <- paste(c(m4$GENE[i], m4$DRUGBANK_NAME[[i]]), collapse=" ")
  #query
  
  res <- EUtilsSummaryCached(query, type="esearch", db="pubmed", retmax=10)
  pmidCnt <- QueryCount(res)
  #pmidCnt
  
  fetch <- EUtilsGetCached(res, type="efetch", db="pubmed")
  pmid <- ArticleId(fetch)
  curPmids <- paste(pmid, collapse = ";")
  
  if(verbose) {
    setTxtProgressBar(pb, i)
  }
  
  pmidCnts <- c(pmidCnts, pmidCnt)
  #pmids <- c(pmids, pmid)
  pmids[[i]] <- pmid
}

#f8 <- cbind(m4, PMID_COUNT=pmidCnts, PMIDS=pmids)
f8 <- m4
f8$PMID_COUNT <- pmidCnts
f8$PMIDS <- pmids

f9 <- convertDataFrameListsToVectors(f8)
#write.table(f9, file="tmp.txt", sep="\t", quote=FALSE, row.names=FALSE)

#f7 <- unique(f9[f9$PMID_COUNT > 0, c("DRUGBANK_CHEBI", "DRUGBANK_NAME")])
#write.table(f7, file="drugbank_blacklist.txt", sep="\t", quote=FALSE, row.names=FALSE)

h1 <- read.table("drugbank_blacklist.txt", sep="\t", header=TRUE, stringsAsFactors = FALSE)
h2 <- h1[which(h1$EXCLUDE), "DRUGBANK_CHEBI"]

f10 <- f9[!(f9$DRUGBANK_CHEBI %in% h2), ]
write.table(f10, file="tmp.txt", sep="\t", quote=FALSE, row.names=FALSE)
```

# Summary
```{r}
# Genes
cat("GENES: Total\n")
length(unique(keggSmMolSif$PARTICIPANT_B))

cat("GENES: Targetable\n")
length(unique(m4$GENE))

# Drugs
cat("DRUGS\n")
length(unique(m4$DRUGBANK_CHEBI))

# Pathways
cat("PATHWAYS\n")
length(sort(unique(unlist(m4$KEGG_PATHWAY_NAMES))))
sort(unique(unlist(m4$KEGG_PATHWAY_NAMES)))
```

## IGNORE: Exclude certain drug types of less interest
```{r, eval=FALSE}
excludeCategories <- c("experimental", "illicit", "nutraceutical", "withdrawn")
c1 <- lapply(drugbankGroupsCategories$DRUGBANK_CATEGORY, function(x) {
  a <- all(!(x %in% excludeCategories))
  b <- length(x) > 0
  r <- a & b
})
c2 <- unlist(c1)
c3 <- cbind(drugbankGroupsCategories, class=c2)
dbOut <- c3[c2,]

# Find the interactions in keggDbSif that match the remaining drugs 
kdIdx <- which(keggDbSif$PARTICIPANT_A %in% dbOut$PARTICIPANT)
kdOut <- keggDbSif[kdIdx, ]

s1 <- merge(kdOut, dbOut, by.x="PARTICIPANT_A", by.y="PARTICIPANT")
drops <- c("INTERACTION_PUBMED_ID", "INTERACTION_DATA_SOURCE", "MEDIATOR_IDS", "class")
# Issues with data.table
setDF(s1)
s1 <- s1[, !(colnames(s1) %in% drops)]

## 
n1 <- convertSifToGmt(keggSmMolSif, "gmt", returnSmallMolecules = FALSE)
genes <- n1$gmt

a1 <- list()

for(i in 1:length(genes)) {
  #i <- 7
  #i <- 62 # COMT
  gene <- genes[i]
  
  kd1 <- filterSif(dbSif$edges, ids=gene, idsBothParticipants = FALSE)
  kd2 <- filterSif(keggSif$edges, ids=gene, interactionTypes = intCats$BetweenProteinSmallMolecule, idsBothParticipants = FALSE)
  
  b2 <- addSifNames(kd1, dbSif$nodes)
  b3 <- addSifNames(kd2, keggSif$nodes)
  b4 <- rbind(b2, b3)
  
  sm1 <- convertSifToGmt(b4, "gmt", returnSmallMolecules = TRUE)
  sm2 <- dbOut$PARTICIPANT %in% sm1$gmt
  sm3 <- NA 
  
  if(any(sm2)) {
    sm3 <- dbOut[sm2, ]
    sm3 <- cbind(sm3, GENE=rep(gene, nrow(sm3)))
    #sm3 <- unique(sm3)
  }
  
  if(!is.na(sm3)) {
    a1[[gene]] <- sm3
  }
}

g1 <- "HPD"
#View(a1$PDE1A)

sm4 <- filterSif(keggSmMolSif, ids = "PDE1A", idsBothParticipants = FALSE)
sm5 <- addSifNames(sm4, keggSif$nodes)
```

## IGNORE: Add SIF Names to Merge Information about the Nodes and Interactions
```{r, eval=FALSE}
addSifNames <- function(tmpSif, tmpNodes) {
  b1 <- merge(tmpSif, tmpNodes, by.x="PARTICIPANT_A", by.y="PARTICIPANT", all.x=TRUE)
  drop <- c("INTERACTION_DATA_SOURCE", "INTERACTION_PUBMED_ID", "PATHWAY_NAMES", "MEDIATOR_IDS", "PARTICIPANT_TYPE", "UNIFICATION_XREF", "RELATIONSHIP_XREF")
  setDF(b1)
  b1 <- b1[, !(colnames(b1) %in% drop)]
  
  b2 <- merge(b1, tmpNodes, by.x="PARTICIPANT_B", by.y="PARTICIPANT", all.x=TRUE)
  drop <- c("PARTICIPANT_TYPE", "UNIFICATION_XREF", "RELATIONSHIP_XREF")
  b2 <- b2[, !(colnames(b2) %in% drop)]
  
  cols <- c("PARTICIPANT_A", "INTERACTION_TYPE", "PARTICIPANT_B", "PARTICIPANT_NAME.x", "PARTICIPANT_NAME.y")
  b2 <- b2[, cols]
  
  return(b2)  
}
```

## Session Info
```{r}
sessionInfo()
```